---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# etlTurtleNesting

<!-- badges: start -->
<!-- badges: end -->

The package etlTurtleNesting contains the ETL and QA for Turtle Nesting Census 
data from ODK Central to WAStD.

Issues are tracked at [wastdr](https://github.com/dbca-wa/wastdr/milestone/1).

## Installation

You can install etlTurtleNesting from [GitHub](https://github.com/) with:

```{r, eval=FALSE, echo=TRUE}
remotes::install_github("dbca-wa/etlTurtleNesting")
```

## Run the WAMTRAM geolocation QA dashboard
### Running the dashboard
Either open "vignettes/qa_wamtram_geolocation.Rmd" and click "Run Document", or run

```{r, eval=FALSE, echo=TRUE}
rmarkdown::run(here::here("vignettes/qa_wamtram_geolocation.Rmd"), shiny_args = list(port = 3838, host = "0.0.0.0"))
```

Follow instructions on the last tab (protocols) to update WAMTRAM through its admin front-end.

### Refresh WAMTRAM data
After a while of updating WAMTRAM, you'll want to see these updates reflected in the dashboard.
To refresh WAMTRAM data, stop the running dashboard, then run the command below in the 
console, restart the R session 
(<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>F10</kbd>) then restart the dashboard 
(<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>K</kbd>).

```{r eval=FALSE, echo=TRUE}
w2_data <- wastdr::download_w2_data(save=here::here("inst/w2.rds"))
```

### Update wastdr
When instructed to upgrade the R package `wastdr` which does most of the work under the bonnet, run:

```{r eval=FALSE, echo=TRUE}
remotes::install_github("dbca-wa/wastdr", dependencies = TRUE, upgrade = "never", build_vignettes = TRUE)
```

## Add new users
New users can be added to WAStD in batches before each field season.
Local coordinators will provide a spreadsheet with columns "name" (full name,
correct! spelling and capitalisation), "email", "phone".

The spreadsheet is post-processed:

* Open with option "quoted columns as text"
* username (A), name (B), email, phone, role (C-E)
* Formula for username: `=SUBSTITUTE(LOWER(B2), " ", "_")`
* Format phone as text and prefix with +61
* Save as CSV with "quote all text columns"

```{r, eval=FALSE}
# Step 1: New users (username, name, phone, email, role)
# HTTP status returned: 400 for existing, 201 for new
# Append new users to spreadsheet: username, name, email, phone, role
library(wastdr)
users <- here::here("inst/users_nin2020.csv") %>%
 readr::read_csv(col_types = "ccccc") %>%
 wastdr::wastd_bulk_post("users", verbose = TRUE)
```


## Import ODKC data, export WAStD data, create reports and outputs

Run `run.R` as a local job.
Full data export and reporting are already part of `run.R`.
