---
title: "WA Sea Turtle Database (WAStD) analysis - Broome special"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
vignette: >
  %\VignetteIndexEntry{wastd}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
params:
  area_name: 'Cable Beach Broome'
editor_options: 
  markdown: 
    wrap: 72
---

```{css, echo=FALSE}
body, .main-container {max-width: 100%; margin-left: auto; margin-right: auto;}
```

```{r setup}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(rmarkdown.html_vignette.check_title = FALSE)
library(wastdr)
# install.packages('leaflet.extras')
library(leaflet.extras)
drake::loadd("wastd_data")
rt <- . %>% reactable::reactable(filterable = T, sortable = T, searchable = T)

x <- wastd_data %>% 
  wastdr::filter_wastd_turtledata(area_name = params$area_name)
```

```{r prepare_export}
zipfn <- wastdr::urlize(params$area_name)
fn <- here::here("inst/reports", zipfn)
if (!fs::file_exists(fn)) {
  "Creating dir {fn}" %>% glue::glue() %>% wastdr::wastdr_msg_info()
  fs::dir_create(fn, recurse = TRUE)
}
```

# Turtle Nests last season 
A eat map of the turtle nests along Cable Beach last season.

* All species
* Season 2020-21
* Location Cable Beach, all sectors

### All nesting data
Including and tallied per area (Broome), seasons (all) and species (all).
This is to get an overview of the data available.

```{r nesting_overview}
x$tracks %>% 
  wastdr::nesting_type_by_area_season_species() %>% 
  rt
```

### Nests of season 2020-21
Only tracks with nests are shown for only the current season (2020-21).
Nests are shown as clusters, which expand when zoomed in or clicked upon.
These clusters are not a heatmap, and resolve down to the invividual nests.

```{r nesting_map_2020}
nests <- x$tracks %>% 
  filter_wastd_season(2020) %>% 
  filter_realspecies() %>% 
  dplyr::filter(nest_type == "successful-crawl")
  
# map2020 <- nests %>% map_tracks(cluster = F, sites = x$sites)
# leaflet.extras::addHeatmap(map2020, data = nests)

nests %>% 
  map_tracks(cluster = F, sites = x$sites) %>%
  leaflet.extras::addHeatmap(data = nests)
```
```{r, fig.height=7}
nests <- x$tracks %>% 
  filter_realspecies() %>% 
  dplyr::filter(nest_type == "successful-crawl") %>% 
  dplyr::mutate(doy = as.numeric(format(datetime, "%j")))

library(ggplot2)
library(ggTimeSeries)

ggplot_calendar_heatmap(nests, "datetime", "latitude") +
  scale_fill_continuous(low = 'green', high = 'red') + 
  facet_wrap(facets = season ~ ., ncol=1)

ggplot(nests, aes(doy, latitude, colour = species)) +
  geom_point() +
  facet_wrap(facets = season ~ ., ncol=1)
```


### Nests of season 2019-20
Only tracks with nests are shown for only the season of 2019-20.
Nests are shown as clusters, which expand when zoomed in or clicked upon.

```{r nesting_map_2019}
x$tracks %>% 
  filter_wastd_season(2019) %>% 
  filter_realspecies() %>% 
  dplyr::filter(nest_type == "successful-crawl") %>% 
  map_tracks(cluster = T, sites = x$sites)
```


### Nests of season 2018-19
Only tracks with nests are shown for only the season of 2018-19.
Nests are shown as clusters, which expand when zoomed in or clicked upon.

```{r nesting_map_2018}
x$tracks %>% 
  filter_wastd_season(2018) %>% 
  filter_realspecies() %>% 
  dplyr::filter(nest_type == "successful-crawl") %>% 
  map_tracks(cluster = T, sites = x$sites)
```


### Nests of season 2017-18
Only tracks with nests are shown for only the season of 2017-18.
Nests are shown as clusters, which expand when zoomed in or clicked upon.

```{r nesting_map_2017}
x$tracks %>% 
  filter_wastd_season(2017) %>% 
  filter_realspecies() %>% 
  dplyr::filter(nest_type == "successful-crawl") %>% 
  map_tracks(cluster = T, sites = x$sites)
```
