---
title: "Turtle morphometrics"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
vignette: >
  %\VignetteIndexEntry{morphometrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  markdown: 
    wrap: 72
params:
  w2_path: 'inst/w2.rds'
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(etlTurtleNesting)
library(wastdr)
library(dplyr)
library(gt)
library(ggplot2)
library(leaflet)
library(leaflet.minicharts)
```

# Data
```{r load_data}
w2 = readRDS(here::here(params$w2_path))
```

The data analysed here is a snapshot of the turtle tagging database retrieved on
`r w2$downloaded_on`.
We exclude data captured at Thevenard and Delambre Island 2021 for now.

# Morphometrics
We show number of observations (n), min, max, mean, sd, and median for morphometric
measurements of type CCL (curved carapace length), CCL Notch (CCL entered into 
wrong field), CCW (curved carapace width) for nesting female Flatback turtles 
by rookery.

Caveats: The dataset could contain incorrect data, mainly with decimal point errors.
The data contain live sightings (daytime in-water captures), night taggings 
of nesting turtles, stranded injured or dead turtles, as well as juvenile animals.

## Morphometrics table

```{r table_data}
morph <- w2$obs_measurements %>% 
  dplyr::left_join(w2$enc, by="observation_id") %>% 
  dplyr::filter(species_code=="FB", nesting=="Y", sex=="F") 

morph_stats <- morph %>% 
  dplyr::group_by(
    location_code,
    measurement_type_code
  ) %>% 
  dplyr::summarise(
    n = dplyr::n(),
    min = min(measurement_value),
    max = max(measurement_value),
    mean = mean(measurement_value),
    sd = sd(measurement_value),
    median = median(measurement_value),
    .groups = "keep"
  ) %>% 
  dplyr::ungroup() %>%
  dplyr::filter(n>100) %>% 
  dplyr::arrange(measurement_type_code, location_code)

morph_stats %>% 
  gt::gt() %>% 
  gt::tab_header(title="Flatback Morphometrics") %>% 
  gt::cols_label(
    location_code = "Rookery",
    measurement_type_code = "Measurement"
  )
```

## Morphometrics histogram 
We show selected morphometrics (CCL, CCL NOTCH, CCW) for the three most sampled
rookeries, Barrow (BW), Dampier Archipelago (DA), Mundabullangana (MN), and
Port Hedland (PH).

```{r plot_data, fig.height=9, fig.width=7}
morph %>% 
  dplyr::filter(
    measurement_type_code %in% c("CCL", "CCL NOTCH", "CCW"),
    location_code %in% c("BW", "DA", "MN", "PH")
    ) %>% 
  dplyr::mutate(
    morphometric = measurement_type_code, 
    measurement = measurement_value) %>% 
  ggplot(aes(x=measurement, fill=morphometric)) +
  ggplot2::labs(
    title="Morphometric measurements of nesting female Flatback turtles",
    subtitle=paste0(
      "Locations: Barrow (BW), Dampier Archipelago (DA), Mundabunllangana (MN), Port Hedland (PH)\n",
      "Measurements: Curved Carapace Length (CCL / CCL NOTCH) and Width (CCW)"),
    x="Measurement [mm]",
    y="Number of turtles"
    ) +
  ggplot2::xlim(c(600, 1000)) +
  geom_histogram(binwidth=5, colour="black") +
  facet_wrap("location_code", ncol=1) +
  theme_classic()
```

