---
title: "WAStD User Mapping QA"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 100%;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(etlTurtleNesting)
library(drake)
library(wastdr)
library(pointblank)
library(magrittr)
library(reactable)
library(gt)
# readd(w2_data)
# readd(user_mapping)
# loadd(w2_data)
# loadd(user_mapping)
```

## Currency

-   This QA report ran on `r Sys.time()`.
-   The ODK Central data was downloaded on `r odkc_ex$downloaded_on`.
-   The WAStD Users were downloaded just before the workbook ran.

## Step 0: Add new data collectors to WAStD

Coordinators of data capture programs supply us with a spreadsheet in this exact
format:

-   Column "name": The full name, including given and surname, e.g. "Florian
    Mayer"

    -   Use correct capitalisation. Do not write "florian mayer" or "FLORIAN
        MAYER". The case does not matter for the user matching, but it matters
        for displaying in reports.

-   Column "email": A valid email, e.g. "<Florian.Mayer@dbca.wa.gov.au>".

    -   Omit any HTML markup, e.g. "mailto:", or anything involving `<`pointy
        brackets`>`.

-   Column "phone": A valid phone number with the Australian prefix `+61`.

    -   Valid example: "+61414123456".
    -   Whitespace will be discarded, so "+61 414 123 456" works too.
    -   Format the column as "text". Excel has a nasty habit of auto-formatting
        columns and dropping the leading "+". This is very tedious to fix. E.g.
        "61414123456" is invalid.

We add columns:

-   "username", generated from the snake-cased "name", e.g. "florian\_mayer".
-   "role", which we set to a predefined text such as "Cable Beach Broome
    Volunteer 2020-21". As user registrations come in batches from distinct
    locations, we get a cleaner result if we set the role centrally following a
    naming convention.

## Step 1: wastdr takes a wild guess

During the data import from ODK Central to WAStD, wastdr maps each distinct ODK
Collect "username" as written by data collectors to actual WAStD user profiles.

The matching is done by
[`fuzzyjoin::stringdist_left_join`](http://varianceexplained.org/fuzzyjoin/reference/stringdist_join.html)
using the Jaro-Winker distance between the ODKC username and the combined WAStD
`name` and `aliases`.

Details about the distance measures are
[here](http://finzi.psych.upenn.edu/library/stringdist/html/stringdist-metrics.html).

The Jaro-Winker distance was chosen as it returns the highest number of correct
matches.

## Step 2: Manually update WAStD user aliases to improve mapping

In the automated validation output we show all user mappings with a
dissimilarity above a threshold. This should separate mismatches from matches.

The validation table shows columns:

-   `odkc_username` - the ODK Collect username
-   `wastd_matched` - a link to the WAStD User profile that we matched
-   `search_wastd` - individual links to search WAStD Users (names, usernames,
    aliases, role, phone, email) for each part of the ODK Collect username. This
    is helpful where a list of names were supplied, e.g. "Flo and Janet".

For each mismatch: Find the correct WAStD user, and add the `odkc_username` to
the WAStD User's field "aliases", then "save" or "save and continue editing" the
WAStD user profile. It is **absolutely essential** to separate the aliases with
a **comma**. e.g. Florian Mayer's aliases can be "Flo Mayer, FlorianM, FloJo",
which then will precisely map to "Flo Mayer", "FlorianM", and "FloJo".

Note: updating WAStD will not refresh this report automatically - we have to
re-run the data import.

When we re-run the user matching with fresh data, this user should match up.
Some will won't, which we'll handle in the next step.

The QA validation will pick up all username matches with a dissimilarity of 0.01
or higher. Perfect matches will have a dissimilarity below 0.01.

You can download a spreadsheet of mismatches from the CSV button:

```{r user_mapping, message = FALSE, echo=FALSE}
a <- user_mapping %>%
    annotate_user_mapping_w2() %>%
    pointblank::create_agent() %>%
    pointblank::col_vals_lt("dist", 0.01) %>%
    interrogate()
a
```

Alternatively, you can work through the issues directly in the table below.

Your actions:

-   Users who put team lists in username - please re-train ASAP:

    -   Username is tablet holder.
    -   SVS team is other names.
    -   Full names as given pre season in Step 0.

-   If a "We matched WAStD User" is definitely not the person from the "ODK
    Collect username":

    -   Find the correct user in WAStD.
    -   Use the individual links in "Search likely candidates" to search WAStD
        for users matching part of the ODK username.
    -   Really not a WAStD user yet? Create new user.
    -   Update the real WAStD user's "aliases" field with the ODK username.
        Important: drop commas from ODK username, but separate distinct aliases
        with a comma. This is crucial for our user matching algorithm.

-   If a "We matched WAStD User" is the person from the "ODK Collect username"
    but has a bad match:

    -   This happens e.g. when the ODK Username is a part (given name), an alias
        (abbreviation), or a team list.

    -   Open the "We matched WAStD user" profile link and update the alias with
        the ODK username. 
    
    -   Next time we run the user matching, the match should come up with a 
        lower dissimilarity measure (or even drop off the validation shortlist).

```{r user_table, echo=FALSE}
validation_data <- 
  a %>%
  pointblank::get_data_extracts() %>%
  magrittr::extract2(1) %>%
  dplyr::rowwise() %>% 
  # dplyr::mutate(active_at = get_user_area(odkc_ex, odkc_username)) %>% 
  dplyr::arrange(
    # active_at, 
    -dist)

# Create missing users in WAStD
# Run manually when necessary
# 
#  missing_users <- validation_data %>% 
#   dplyr::select(legacy_userid) %>% 
#   dplyr::arrange(legacy_userid) %>%
#   dplyr::distinct(legacy_userid) %>% 
#   dplyr::left_join(w2_data$persons, by = c("legacy_userid"="person_id")) %>% 
#   dplyr::transmute(
#     username = clean_name %>% wastdr::urlize() %>% stringr::str_replace_all("-", "_"),
#     name = clean_name,
#     # phone = mobile, # NA
#     # email = clean_email, # NA
#     role = glue::glue("{specialty} {comments}")
#   )
#  missing_users %>% wastdr::wastd_bulk_post("users", verbose = TRUE)

validation_data %>% 
  gt::gt() %>%
  gt::fmt_markdown(columns = TRUE) %>%
  gt::cols_label(
    legacy_username = gt::html("<h5>They wrote</h5>\n<small>ODK Collect username</small>"),
    # active_at = gt::html("<small>Active at</small>"),
    legacy_un_trim = gt::html("<h5>We searched</h5>\n<small>Cleaned ODK name</small>"),
    wastd_matched = gt::html("<h5>We matched</h5>\n<small>WAStD User</small>"),
    search_wastd = gt::html("<h5>Search</h5>\n<small>likely candidates</small>"),
    dist = gt::html("<h5>Dissimilarity</h5>\n<small>smaller = better</small>")
  ) %>%
  gt::tab_spanner(
    label = "WASTD User match",
    columns = vars(legacy_username,
                   # active_at,
                   legacy_un_trim,
                   wastd_matched, 
                   search_wastd, 
                   dist)
  ) %>%
  gt::tab_spanner(
    label = "WASTD User profile (chosen as most likely)",
    columns = vars(role, pk, username, name, nickname, aliases, email, phone)
  )
```
