---
title: "WAStD User Mapping QA"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 100%;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(etlTurtleNesting)
library(drake)
library(wastdr)
library(pointblank)
library(magrittr)
library(reactable)
library(gt)
# readd(w2_data)
# readd(w2_user_mapping)
# loadd(w2_data)
# loadd(w2_user_mapping)
```


```{r user_table, echo=FALSE, message=FALSE}
a <- user_mapping %>%
    annotate_user_mapping_w2() %>%
    pointblank::create_agent() %>%
    pointblank::col_vals_lt("dist", 0.01) %>%
    interrogate()

if (!is.null(a$has_intel)){
actions_table <- a %>%
  pointblank::get_data_extracts() %>%
  magrittr::extract2(1) %>%
  dplyr::rowwise() %>% 
  # dplyr::mutate(active_at = get_user_area(odkc_ex, odkc_username)) %>% 
  # dplyr::select(-email, -phone) %>% 
  dplyr::arrange(-dist) %>% 
  gt::gt() %>%
  gt::fmt_markdown(columns = gt::everything()) %>%
  gt::cols_label(
    legacy_username = gt::html("<h5>They wrote</h5>\n<small>W2 username</small>"),
    legacy_userid = gt::html("<small>W2 legacy ID</small>"),
    # active_at = gt::html("<small>Active at</small>"),
    legacy_un_trim = gt::html("<h5>We searched</h5>\n<small>Cleaned W2 name</small>"),
    wastd_matched = gt::html("<h5>We matched</h5>\n<small>WAStD User</small>"),
    search_wastd = gt::html("<h5>Search</h5>\n<small>likely candidates</small>"),
    dist = gt::html("<h5>Dissimilarity</h5>\n<small>smaller = better</small>")
  ) %>%
  gt::tab_spanner(label = "Wamtram profile",
                  columns = c(legacy_username,
                              legacy_userid)) %>%
  gt::tab_spanner(label = "WASTD User match",
                  columns = c(legacy_un_trim,
                              wastd_matched,
                              search_wastd,
                              dist)) %>%
  gt::tab_spanner(
    label = "WASTD User profile (chosen as most likely)",
    columns = c(role, pk, username, name, nickname, aliases, email, phone)
  )  
} else {actions_table <- NULL}
```

# User matching results {.tabset}

## Please review {.active}
Work through the issues directly in the table below.

Your actions:

-   Users who put team lists in username - please re-train ASAP:

    -   Username is tablet holder.
    -   SVS team is other names.
    -   Full names as given to us pre season.

-   If a "We matched WAStD User" is definitely not the person from the "ODK
    Collect username":

    -   Find the correct user in WAStD.
    -   Use the individual links in "Search likely candidates" to search WAStD
        for users matching part of the ODK username.
    -   Really not a WAStD user yet? Create new WAStD user.
    -   Update the real WAStD user's "aliases" field with the ODK username.
        Important: drop commas from ODK username, but separate distinct aliases
        with a comma. This is crucial for our user matching algorithm.

-   If a "We matched WAStD User" is the person from the "ODK Collect username"
    but has a bad match:

    -   This happens e.g. when the ODK Username is a part (given name), an alias
        (abbreviation), or a team list.

    -   Open the "We matched WAStD user" profile link and update the alias with
        the ODK username. 
    
    -   Next time we run the user matching, the match should come up with a 
        lower dissimilarity measure (or even drop off the validation shortlist).

```{r view_match_table, echo=FALSE}
if (is.null(actions_table)) {
  wastdr::wastdr_msg_success("No mismatches found")
} else {
    actions_table
}
```


## Data and Methods

-   This QA report ran on `r Sys.time()`.
-   The ODK Central data was downloaded on `r w2_data$downloaded_on`.
-   The WAStD Users were downloaded just before the workbook ran.

During the data import from ODK Central to WAStD, wastdr maps each distinct ODK
Collect "username" as written by data collectors to actual WAStD user profiles.

The matching is done by
[`fuzzyjoin::stringdist_left_join`](http://varianceexplained.org/fuzzyjoin/reference/stringdist_join.html)
using the Jaro-Winker distance between the ODKC username and the WAStD
`name` and `aliases`. Aliases are split up by comma. 
The Jaro-Winker distance was chosen as it returns the highest number of correct
matches.

Details about available distance measures can be found
[here](http://finzi.psych.upenn.edu/library/stringdist/html/stringdist-metrics.html).


You can download a spreadsheet of mismatches from the CSV button:

```{r user_mapping, message = FALSE, echo=FALSE}
a
```

Note: updating WAStD will not refresh this report automatically - we have to
re-run the data import.

When we re-run the user matching with fresh data, users with updated aliases 
should match up.

The QA validation will pick up all username matches with a dissimilarity of 0.01
or higher. Perfect matches will have a dissimilarity below 0.01.

## Adding new data collectors to WAStD

Coordinators of data capture programs supply us with a spreadsheet in this exact
format:

-   Column "name": The full name, including given and surname, e.g. "Florian
    Mayer"

    -   Use correct capitalisation. Do not write "florian mayer" or "FLORIAN
        MAYER". The case does not matter for the user matching, but it matters
        for displaying in reports.

-   Column "email": A valid email, e.g. `Florian.Mayer@dbca.wa.gov.au`.

    -   Omit any HTML markup, e.g. "mailto:", or anything involving `<`pointy
        brackets`>`.

-   Column "phone": A valid phone number with the Australian prefix `+61`.

    -   Valid example: "+61414123456".
    -   Whitespace will be discarded, so "+61 414 123 456" works too.
    -   Format the column as "text". Excel has a nasty habit of auto-formatting
        columns and dropping the leading "+". This is very tedious to fix. E.g.
        "61414123456" is invalid.

We add the following columns:

-   "username", generated from the snake-cased "name", e.g. `florian_mayer`.
-   "role", which we set to a predefined text such as "Cable Beach Broome
    Volunteer 2020-21". As user registrations come in batches from distinct
    locations, we get a cleaner result if we set the role centrally following a
    naming convention. Less work for the coordinator!

## Create missing users in WAStD
Run manually when necessary.

```{r, eval=FALSE}
validation_data_extracts <- a %>%
  pointblank::get_data_extracts() 

if (length(validation_data_extracts) > 0) {
  missing_users <- validation_data_extracts %>%
    magrittr::extract2(1) %>% 
    dplyr::rowwise() %>%
    dplyr::select(legacy_userid) %>%
    dplyr::arrange(legacy_userid) %>%
    dplyr::distinct(legacy_userid) %>%
    dplyr::left_join(w2_data$persons, by = c("legacy_userid" = "person_id")) %>%
    dplyr::transmute(
      username = clean_name %>%
        wastdr::urlize() %>% stringr::str_replace_all("-", "_"),
      name = clean_name,
      # phone = mobile, # NA
      # email = clean_email, # NA
      role = glue::glue("{specialty} {comments}")
    )
  
  missing_users %>% wastdr::wastd_bulk_post("users", verbose = TRUE)
  
}
```
