---
title: "WAMTRAM QA"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---

```{css, echo=FALSE}
#map {height: calc(100vh - 80px) !important;}
```

```{r global, include=FALSE}
# Libraries -------------------------------------------------------------------#
library(wastdr)
library(shiny)
library(flexdashboard)
library(leaflet)
library(fs)
library(here)
library(dplyr)
library(tidyr)
library(geojsonio)
library(geojsonsf)
library(janitor)
library(reactable)
library(RODBC)
library(DBI)

# Settings --------------------------------------------------------------------#
knitr::opts_chunk$set(echo = FALSE)

# Data ------------------------------------------------------------------------#
#
# WAMTRAM
w2_dir <- fs::dir_create(here::here("inst"))
w2_file <- here::here("inst/w2.rds")
if (fs::file_exists(w2_file)){ 
  wastdr::wastdr_msg_info("Reading saved WAMTRAM data...")
  w2_data <- readRDS(w2_file)
} else {
  wastdr::wastdr_msg_info("Downloading WAMTRAM data...")
  w2_data <- wastdr::download_w2_data(save=w2_file)
}
locations <- c("", sort(unique(w2_data$enc$location_code)))
places <- c("", sort(unique(w2_data$enc$place_code)))
w2_dl <- w2_data$downloaded_on

wastdr::wastdr_msg_info("Downloading WAStD Areas...")
areas_sf <- wastdr::wastd_GET("area") %>%
  magrittr::extract2("data") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf()

areas_dl <- Sys.time()

areas <- areas_sf %>%
  dplyr::filter(area_type == "Locality") %>%
  dplyr::transmute(area_id = pk,
                   area_name = name,
                   w2_location_code = w2_location_code)

sites <- areas_sf %>%
  dplyr::filter(area_type == "Site") %>%
  dplyr::transmute(site_id = pk,
                   site_name = name,
                   w2_place_code = w2_place_code) %>%
  sf::st_join(areas)

sites_by_pc <- sites %>% 
  tidyr::separate_rows(w2_place_code)

enc_by_sites <- w2_data$enc %>% 
  dplyr::group_by(place_code) %>% 
  dplyr::tally(name = "observations") %>% 
  dplyr::ungroup()

homeless_places <- w2_data$sites %>% 
  dplyr::filter(is.na(site_latitude)) %>% 
  dplyr::mutate(search_areas_at=glue::glue(
    '<a href="https://wastd.dbca.wa.gov.au/admin/observations/area/?q={prefix}"',
    ' target="_" rel="external"><strong>{prefix}</strong></a>'),
    .before=code) %>% 
  dplyr::select(-site_datum, site_longitude, site_latitude) %>% 
  dplyr::left_join(enc_by_sites, by=c("code"="place_code")) %>% 
  dplyr::filter(observations > 0) %>% 
  dplyr::arrange(-observations) %>% 
  janitor::clean_names(case="title")

located_places <- w2_data$sites %>% 
  dplyr::filter(!is.na(site_latitude)) %>% 
  dplyr::mutate(search_areas_at=glue::glue(
    '<a href="https://wastd.dbca.wa.gov.au/admin/observations/area/?q={prefix}"',
    ' target="_" rel="external"><strong>{prefix}</strong></a>'),
    .before=code) %>% 
  dplyr::select(-site_datum, site_longitude, site_latitude) %>% 
  # dplyr::left_join(sites_by_pc, by=c("code"="w2_location_code")) %>% 
  dplyr::left_join(enc_by_sites, by=c("code"="place_code")) %>% 
  dplyr::filter(observations > 0) %>%
  dplyr::arrange(-observations) %>% 
  janitor::clean_names(case="title")

invalid_coords <- w2_data$enc %>% 
  dplyr::filter(
    longitude < -180 | 
      longitude > 180 | 
      latitude < -90 | 
      latitude > 90 |
      is.na(latitude) |
      is.na(longitude)
      )

unlikely_coords <- w2_data$enc %>% 
  dplyr::filter(
    longitude < 100 | longitude > 150 | latitude < -45 | latitude > 0)

w2_obs_wastd_sites <- w2_data$enc %>%
  dplyr::filter(
    longitude > -180, longitude < 180, latitude > -90, latitude < 90) %>%
  sf::st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  sf::st_join(sites, left = TRUE)
```

Places
================================================================================
```{r map_loc, echo=FALSE}
renderLeaflet({
  wastdr::map_wastd_wamtram_sites(areas, sites, w2_data$sites)
})
```

Observations {data-icon="fa-map-location-dot"}
================================================================================

Column {.sidebar}
--------------------------------------------------------------------------------
### Filter Observations

```{r input_obs}
inputPanel(
  selectInput("lc", label = "Location:", choices = locations),
  selectInput("pc", label = "Place:", choices = places),
  textInput("oid", label = "Observation ID:", value = "")
)
```

### Place selected

```{r place_coords}
selected_place <- reactive({w2_data$sites %>% dplyr::filter(code == input$pc)})
renderText({glue::glue("Lat {round(selected_place()$site_latitude, 5)}")})
renderText({glue::glue("Lon {round(selected_place()$site_longitude, 5)}")})
```

Column {data-width=150}
--------------------------------------------------------------------------------
### WAMTRAM

```{r gauge_w2_dl}
valueBox(w2_dl, icon = "fa-calendar", caption = "WAMTRAM")
```

### WAStD 

```{r gauge_area_dl}
valueBox(areas_dl, icon = "fa-calendar", caption = "WAStD Areas")
```

### Total Obs

```{r}
total_obs <- reactive({
  ifelse(
    (is.null(input$pc) || input$pc == ""),
    w2_obs_wastd_sites %>% nrow(),
    w2_obs_wastd_sites %>% dplyr::filter(place_code == input$pc) %>% nrow()
  )
})
# renderText({total_obs()})
renderValueBox({
  valueBox(total_obs(), icon = "fa-map", caption = "Total Obs")
})
```

### Obs inside sites

```{r}
total_obs_inside <- reactive({
  ifelse(
    (is.null(input$pc) || input$pc == ""),
    w2_obs_wastd_sites %>% dplyr::filter(!is.na(site_name)) %>% nrow(),
    w2_obs_wastd_sites %>%
      dplyr::filter(place_code == input$pc, !is.na(site_name)) %>% nrow()
  )
})
renderValueBox({
  valueBox(total_obs_inside(), icon = "fa-map-marker", caption = "Obs in Sites")
})
```


### Obs outside sites

```{r}
total_obs_outside <- reactive({
  ifelse(
    (is.null(input$pc) || input$pc == ""),
    w2_obs_wastd_sites %>% dplyr::filter(is.na(site_name)) %>% nrow(),
    w2_obs_wastd_sites %>%
      dplyr::filter(place_code == input$pc, is.na(site_name)) %>% nrow()
  )
})
renderValueBox({
  valueBox(total_obs_outside(), 
           icon = "fa-question", 
           caption = "Obs not in Sites")
})
```


### Pct inside sites

```{r}
pct_obs <- reactive({round(100 * total_obs_inside() / total_obs(), 1)})
renderValueBox({
  
  valueBox(pct_obs(), 
           icon = "fa-percent", 
           caption = "Pct in Sites",
           color = ifelse(pct_obs() > 95, "success", "warning"))
})
```


Column
--------------------------------------------------------------------------------
### Map
```{r map_obs}
renderLeaflet({
  wastdr::map_wamtram(
    w2_data, 
    location = input$lc, 
    place = input$pc, 
    obs_id = input$oid, 
    wa_sites = sites,
    l_height="calc(100vh - 80px)"
  )
})
```

Bad coords {data-icon="fa-danger"}
================================================================================
### Impossible coordinates
```{r}
reactable::renderReactable({
  reactable::reactable(
    invalid_coords, 
    sortable = TRUE, 
    filterable = TRUE, 
    searchable = TRUE,
    defaultColDef = colDef(html = TRUE)
    )
})
```


### Unlikely coordinates
```{r}
reactable::renderReactable({
  reactable::reactable(
    unlikely_coords, 
    sortable = TRUE, 
    filterable = TRUE, 
    searchable = TRUE,
    defaultColDef = colDef(html = TRUE)
    )
})
```


Places with Homes &trade;
================================================================================
### Place statistics
```{r}
renderValueBox({
  place_loc_rate <- round(100*(nrow(located_places) / nrow(w2_data$sites)), 0)
  flexdashboard::valueBox(
    value = place_loc_rate,
    caption = "% located",
    icon = "fa-area-chart",
    color = if (place_loc_rate <= 90) "warning" else "primary"
  )
})
```


### Create WAStD sites for these WAMTRAM places
```{r}
reactable::renderReactable({
  reactable::reactable(
    located_places, 
    sortable = TRUE, 
    filterable = TRUE, 
    searchable = TRUE,
    defaultColDef = colDef(html = TRUE)
    )
})
```

Places without Homes &trade;
================================================================================
### Place statistics
```{r}
renderValueBox({
  place_loc_rate <- round(100*(nrow(homeless_places) / nrow(w2_data$sites)), 0)
  flexdashboard::valueBox(
    value = place_loc_rate,
    caption = "% missing coordinates",
    icon = "fa-area-chart",
    color = if (place_loc_rate <= 90) "warning" else "primary"
  )
})
```

### Update WAMTRAM with coordinates for these places
```{r}
reactable::renderReactable({
  reactable::reactable(
    homeless_places, 
    sortable = TRUE, 
    filterable = TRUE, 
    searchable = TRUE,
    defaultColDef = colDef(html = TRUE)
    )
})
```

Protocols
================================================================================

### Places
* Purpose: WAStD Areas of type Locality (WAMTRAM location) and Site (WAMTRAM Place) 
  should correspond to all WAMTRAM Places within their bounds.
* WAStD Sites should be one nesting beach that is typically surveyed in one effort.
  Some nesting beaches are so large that they are split into sectors, which
  are surveyed individually.
* Some WAStD Sites can or should contain multiple WAMTRAM Places.
* We can split WAStD Sites later and re-link the WAMTRAM Places.


Red place markers are WAMTRAM sites without corresponding WAStD sites.
For each red marker:

* Create a WAStD Area of type "Site".
  * Create a polygon that buffers one nesting beach by ca 100m, include vegetation
    and shallow water.
  * If there are other Sites close by, enable "WAStD Sites" from the layer menu.
  * Go fullscreen for more convenient digitising if needed.
  * Populate `w2_location_code` with exactly one W2 location code, taken from the 
    red W2 Place marker. Code is all caps.
  * Populate `w2_place_code` with any W2 Place's codes that are contained within.
    Separate multiple values with a whitespace. Codes are all caps.
  * OK to ignore "survey length" and "roundtrip length" for now.
* If we ever need to report on surveys to a bunch of neighbouring Sites, 
  create a WAStD Area with the Place's location code as `w2_location_code`. 
  The Area should contain all Sites within.
  
WAStD Areas (Localities and Sites) are downloaded every time this dashboard is run.

### Observations

* Purpose: Fix errors in W2 Observation coordinates.
* Select one place code. 
* Review coordinates of misplaced observations (popup) against the coordinates 
  of the selected place (sidebar).
* Most typos are immediately obvious, as the first two to three decimals of each
  coordinate should be the same or close for place and observation.
* Find observations through their observation ID in WAMTRAM and fix their coordinates.
* Leave a note in the observation comments about the coordinate change, e.g.
  "Lon 11123.456 changed to 123.456". This is the only change history we have.

To refresh WAMTRAM data, stop the dashboard, run the command below in the 
console, restart the R session 
(<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>F10</kbd>) then restart the dashboard 
(<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>K</kbd>).

```{r eval=FALSE, echo=TRUE}
w2_data <- wastdr::download_w2_data(save=here::here("inst/w2.rds"))
```

To load new `wastdr` functionality, run (when instructed to upgrade `wastdr`):

```{r eval=FALSE, echo=TRUE}
remotes::install_github("dbca-wa/wastdr")
```

### Places without Home &trade;
Search WAStD for places at the same location code, create a WAStD Site for the 
WAMTRAM place with the missing coordinates, save, then use the centroid you can
see in the WAStD Area list view as coordinates in WAMTRAM.
