---
title: "Strandings 2016-18"
author: "Florian Mayer, DBCA"
date: "27/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(wastdr)
```

```{r load_data}
drake::loadd("wastd_data")

rt <- . %>% 
  reactable::reactable(
    ., 
    filterable = T, 
    sortable = T, 
    searchable = T,
    defaultColDef = reactable::colDef(html = TRUE)
  )
```

Data was accessed on `r wastd_data$downloaded_on`.

We filter AnimalEncounters to those with turtles in calendar years 2016-18
and split them by live and dead turtles.
We exclude tagged turtles (which can also record damages/injuries).

```{r filter_data}
ani <- wastd_data$animals %>% 
  dplyr::mutate(calendar_year = datetime %>% lubridate::year()) %>% 
  dplyr::filter(calendar_year>=2016, calendar_year<=2018) %>%  
  dplyr::filter(taxon=="Cheloniidae")

ani_dead <- ani %>% filter_dead()
ani_alive <- ani %>% filter_alive() %>% dplyr::filter(health!="na")

dmg <- wastd_data$turtle_dmg %>% 
  dplyr::filter(calendar_year>=2016, calendar_year<=2018)
morph <- wastd_data$turtle_morph %>% 
  dplyr::filter(calendar_year>=2016, calendar_year<=2018)

```

# Exports
Both alive and dead turtles are exported to CSV and RDS (to open in R), 
plus damages and morphometrics for combined dead and alive animals.
```{r export_data}
dest <- "inst/reports/strandings_2016_2018"

ani_alive %>% saveRDS(here::here(dest, "alive.rds"))
ani_alive %>% readr::write_csv(here::here(dest, "alive.csv"))

ani_dead %>% saveRDS(here::here(dest, "dead.rds"))
ani_dead %>% readr::write_csv(here::here(dest, "dead.csv"))

dmg %>% saveRDS(here::here(dest, "dmg.rds"))
dmg %>% readr::write_csv(here::here(dest, "dmg.csv"))

morph %>% saveRDS(here::here(dest, "morph.rds"))
morph %>% readr::write_csv(here::here(dest, "morph.csv"))
```


# Strandings
This map provides a quick overview of stranding locations.

```{r map_dead}
ani_dead %>% map_mwi()
```
The following table previews the contents of the CSV "ani_dead.csv".

```{r show_dead}
ani_dead %>% rt()
```


# Rescues / non-mortalities

This map provides a quick overview of rescue/non-mortalities locations.

```{r map_alive}
ani_alive %>% map_mwi()
```


The following table previews the contents of the CSV "ani_alive.csv".

```{r show_alive}
ani_alive %>% rt()
```

# Damages (dead and alive)
```{r show_dmg}
dmg %>% rt
```

# Morphometrics (dead and alive)
```{r show_morph}
morph %>% rt
```

