---
title: "WA Sea Turtle Database (WAStD) analysis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
vignette: >
  %\VignetteIndexEntry{wastd}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
params:
  area_name: 'Ningaloo'
editor_options: 
  markdown: 
    wrap: 72
---

```{css, echo=FALSE}
body, .main-container {max-width: 100%; margin-left: auto; margin-right: auto;}
```

```{r setup}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(rmarkdown.html_vignette.check_title = FALSE)
library(wastdr)
drake::loadd("wastd_data")
rt <- . %>% 
  reactable::reactable(
    ., 
    filterable = T, 
    sortable = T, 
    searchable = T,
    defaultColDef = colDef(html = TRUE)
  )
```

```{r prepare_export}
zipfn <- wastdr::urlize(params$area_name)
fn <- here::here("inst/reports", zipfn)
if (!fs::file_exists(fn)) {
  "Creating dir {fn}" %>% glue::glue() %>% wastdr::wastdr_msg_info()
  fs::dir_create(fn, recurse = TRUE)
}
```

```{r filter_data}
x <- wastd_data %>% 
  wastdr::filter_wastd_turtledata(area_name = params$area_name)
```

# Overview
This top level summary is exported as `just_the_data_thanks.csv` and
shows totals of tracks, disturbance / predation, live sightings and strandings 
plus track tallies grouped and sorted per area, site, and calendar date (AWST).
Individually encountered tracks are split up by nesting success. 
Track tallies, which are only collected in specific places like Cape Domett,
are counted but the number of tracks therein is not further expanded.

A very detailed explanation of the columns in this table is hosted 
[here]().

```{r top_level_summary}
summarise_wastd_data_per_day_site(x) %T>% 
  readr::write_csv(here::here(fn, "just_the_data_thanks.csv")) %>% 
  rt
```



This report is a story **about** data. It explains, visualises, and
rudimentarily summarises the data which is exported to CSV and various
other formats. The audience of this report are data analysts and
modellers who need to consume and understand the exported data.
Suggestions and ideas for additional summaries or visualisations are
welcome [here](https://github.com/dbca-wa/wastdr/issues).

```{r export_data}
x %>% wastdr::export_wastd_turtledata(outdir = fn, filename = zipfn)
# readRDS(fs::path(fn, "wastd_data.rds")) to restore x
```

# Data

Data for this summary is supplied from WAStD, and is manipulated using
helpers from the R package [wastdr](https://dbca-wa.github.io/wastdr/).
Analysts are welcome to utilise `wastdr` or contribute ideas back as
feature requests.

The raw data is exported into separate CSV files. The content of the
individual CSV files is documented at
[wastdr::export_wastd_turtledata](https://dbca-wa.github.io/wastdr/reference/export_wastd_turtledata.html).

Additionally, an R object of class `wastd_data` containing all the data
objects is saved to `wastd_data.RData` which can be read into R directly.
All code in this workbook can be reproduced by running the code shown
inline with the data from `wastd_data.RData`.

In this report, we filter the data to the location `r params$area_name`.

The data was downloaded from WAStD on `r wastd_data$downloaded_on`.

# Data summaries
Training and test data is marked as species "Hatchback turtle (Corolla corolla)",
surveys are "production=FALSE", disturbance causes are "training or test record".
Where summaries hide the columns "species" and "production", respectively, 
training and test data are excluded.

## Turtle tracks (current season)

```{r tracks_map_2020}
x$tracks %>% 
  wastdr::filter_wastd_season(2020) %>% 
  wastdr::map_tracks(sites = x$sites)
```

### Missed tracks

This section is relevant for monitoring programs which tag turtles.

Missed tracks come from turtles which got away just before or during
tagging. These records can be used to inform tagging start and end
times:

Missed tracks recorded at the beginning of a tagging night means that
tagging needs to start earlier in the day. Missed tracks recorded during
tagging indicates that tagging beaches are understaffed.

Missed tracks are contained in the exported `tracks.csv` with `nest_age` "missed".

```{r tracks_missed}
mt <- x$tracks %>% 
  wastdr::filter_wastd_season(2020) %>% 
  dplyr::filter(nest_age == "missed")

# mt %>% rt
mt %>% wastdr::map_tracks(sites = x$sites)
```

## Disturbance, predation: nests and general
General disturbance and signs of predator presence is captured through the 
ODK form "Predator or Disturbance". Dist urbance or predation of nests are 
captured through the ODK form "Turtle Track or Nest".
As the data specific to the disturbance or predation is identical, both are 
presented here as one table.

In the exported `nest_dist.csv`, the `encounter_type` "nest" refers to disturbances
or predation of turtle nests, whereas "other" refers to sightings of general
disturbance or presence of predators.

```{r dist_map_2020}
x$nest_dist %>% 
  wastdr::filter_wastd_season(2020) %>%
  wastdr::map_dist()
```

## Turtle tracks and nests season totals

Tracks and nests are tallied by type and grouped by

-   season: numeric year of season start, i.e. 2019 = season 2019-20

-   nest age:

    -   fresh - made last night,
    -   old - made before last night,
    -   missed - at night during tagging
    
-   species: "cheloniidae-fam" = species unknown

To see only missed tracks, type "mis" into the filter box of column
"(details\_)nest_age".

This summary is exported as `nesting_type_by_season_age_species.csv`.

```{r tracks_season}
wastd_nesting_type_by_season_age_species <-
  x$tracks %>% 
  wastdr::nesting_type_by_area_season_age_species() %>% 
  dplyr::arrange(-season)

wastd_nesting_type_by_season_age_species %>%
  readr::write_csv(here::here(fn, "nesting_type_by_season_age_species.csv"))
  
wastd_nesting_type_by_season_age_species %>% rt
```

## Turtle tracks and nests daily totals

The same data, but grouped by day instead of season.
This summary is exported as `nesting_type_by_day_age_species.csv`.

```{r tracks_day}
wastd_nesting_type_by_day_age_species <-
  x$tracks %>% 
  wastdr::nesting_type_by_season_calendarday_age_species() %>% 
  dplyr::arrange(-season)

wastd_nesting_type_by_day_age_species %>%
  readr::write_csv(here::here(fn, "nesting_type_by_day_age_species.csv"))
  
wastd_nesting_type_by_day_age_species %>% rt
```

## Night recaptures by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual new turtles by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual recaptures by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual total turtles (individuals) by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual total turtle emergences

Tracks and nests are combined and tallied, grouped by season and species
to form the total number of emergences in one season. This data needs to
be added to the number of tagged turtles from the tagging database.

```{r total_emerg}
x$tracks %>% 
  wastdr::exclude_training_species() %>% 
  dplyr::group_by(season, species) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season) %>%
  rt
```

## Number of nests tagged

Tagged nests are exported as individual records of encounters with
tagged nests (including initial tagging and re-sighting). 
We show the season tallies, and export the individual records as 
`nest_tags.csv`.

```{r nests_tagged}
wastd_tagged_nests <- x$nest_tags %>% 
  wastdr::exclude_training_species()
```

```{r}
# Season tally
# wastd_tagged_nests %>% rt

wastd_tagged_nests %>% 
  dplyr::group_by(season, species) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season) %>%
  rt
```

## Number of nests found

The data here is the same as above, nest type "successful-crawl"
(confirmed fresh nest) and "hatched nest".

This summary is exported as `fresh_and_hatched_nests.csv`.

```{r nests_total}
wastd_nests <- x$tracks %>% 
  wastdr::exclude_training_species() %>% 
  dplyr::filter(nest_type %in% c("successful-crawl", "hatched nest")) %>% 
  dplyr::group_by(season, species, nest_type) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season)

wastd_nests %>%
  readr::write_csv(here::here(fn, "fresh_and_hatched_nests.csv"))

wastd_nests %>% rt
```

## Nest loggers

Nest loggers, n is number of encounters which includes both initial
deployment and any recorded re-sighting.

For want of a better overview, we show an interactive table of all loggers here.
The nest logger records are also exported as `nest_loggers.csv`.

```{r}
x$nest_loggers %>% rt
```

## Hatching and emergence success

Data shown here is summarised by season, and exported as individual nest
excavation records. Data is split up by source between new season
(2019-20) and previous seasons.

This summary is exported as `nest_excavations_hatching_emergence_success.csv`.

```{r nest_excav}
nest_excav <- x$nest_excavations %>%
  wastdr::hatching_emergence_success() %>% 
  dplyr::arrange(-season)

nest_excav %>% 
  readr::write_csv(
    here::here(fn, "nest_excavations_hatching_emergence_success.csv"))

nest_excav %>% rt
```

<!-- TODO Plot HS/ES time series to show different season length / survey effort. -->

## Fan angle data

Fan angle data is exported as `nest_fans.csv` and previewed in the following table.
More meaningful visualisations are in preparation.

```{r fanangle_table}
x$nest_fans %>% rt
```

## Tag return crossovers from or to by year

This section will be completed once the tagging database has been merged into WAStD.

## Strandings, rescues, and live sightings

### Live sightings: rescues, in water, tagging
The ODK form "Turtle Sighthing" captures turtles encountered during boat based 
surveys.
The ODK form "Marine Wildlife Incident" also captures encounters with injured,
entangled, rescued, but overall alive animals (including turtles).

Both forms feed into WAStD AnimalEncounters, which are shown in the map below.

```{r mwi_map_resc}
x$animals %>% 
  wastdr::filter_alive() %>% 
  wastdr::map_mwi(sites = x$sites)
```

### Dead sightings: strandings, mortalities
The ODK form "Marine Wildlife Incident" captures encounters with dead animals 
(including turtles), which then become WAStD AnimalEncounters.

```{r mwi_map_str}
x$animals %>% 
  wastdr::filter_dead() %>% 
  wastdr::map_mwi(sites = x$sites)
```

## Survey effort

A summary of past seasons follows. Note the dates are "turtle dates"
(calendar date of nesting night's evening).

This summary is exported as `survey_season_stats.csv`.

```{r svy_stats, fig.width=10}
survey_season_stats <- x$surveys %>% 
  wastdr::survey_season_stats() 

survey_season_stats %>% 
  readr::write_csv(here::here(fn, "survey_season_stats.csv"))

survey_season_stats %>% rt
```

## QA: Duplicate production surveys

The following table shows days and sites for which more than one production survey
exists in WAStD.
Typically, only one of them should be a "production" survey, 
and all other surveys could be marked and closed as duplicates.

```{r}
x$surveys %>% wastdr::duplicate_surveys() %>% rt
```

## Number of volunteers each year

The number of surveys lead by each volunteer is tallied here for each
season.

```{r volunteers}
x$surveys %>% 
  dplyr::group_by(season, reporter) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season) %>%
  rt
```

## Number of Indigenous employee days

<!-- Scott -->
To be inserted.

## List of sat tags applied in last three years

Source: Seaturtle.org or tagging database.
