---
title: "WA Sea Turtle Database (WAStD) analysis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
vignette: >
  %\VignetteIndexEntry{wastd}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
params:
  area_name: 'Ningaloo'
editor_options: 
  markdown: 
    wrap: 72
---

```{css, echo=FALSE}
body, .main-container {max-width: 100%; margin-left: auto; margin-right: auto;}
```

```{r setup}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(rmarkdown.html_vignette.check_title = FALSE)
library(wastdr)
library(sparkline)
drake::loadd("wastd_data")

x <- wastd_data %>% 
  wastdr::filter_wastd_turtledata(area_name = params$area_name)

area_id <- wastd_data$sites %>% 
  dplyr::filter(area_name==params$area_name) %>% 
  magrittr::extract2("area_id") %>% 
  magrittr::extract(1)

rt <- . %>% 
  reactable::reactable(
    ., 
    filterable = T, 
    sortable = T, 
    searchable = T,
    defaultColDef = reactable::colDef(html = TRUE)
  )

zipfn <- wastdr::urlize(params$area_name)
fn <- here::here("inst/reports", zipfn)
if (!fs::file_exists(fn)) {
  "Creating dir {fn}" %>% glue::glue() %>% wastdr::wastdr_msg_info()
  fs::dir_create(fn, recurse = TRUE)
}
```

# Overview
This top level summary is exported as `just_the_data_thanks.csv` and
shows totals of tracks, disturbance / predation, live sightings and strandings 
plus track tallies grouped and sorted per area, site, and calendar date (AWST).
Individually encountered tracks are split up by nesting success. 
Track tallies, which are only collected in specific places like Cape Domett,
are counted but the number of tracks therein is not further expanded.

A very detailed explanation of the columns in this table is hosted 
[here](https://dbca-wa.github.io/wastdr/reference/summarise_wastd_data_per_day_site.html).

```{r top_level_summary}
summarise_wastd_data_per_day_site(x) %T>% 
  readr::write_csv(here::here(fn, "just_the_data_thanks.csv")) %>% 
  janitor::clean_names(case="sentence") %>%
  rt
```

This report is a story **about** data. It explains, visualises, and
rudimentarily summarises the data which is exported to CSV and various
other formats. The audience of this report are data analysts and
modellers who need to consume and understand the exported data.
Suggestions and ideas for additional summaries or visualisations are
welcome [here](https://github.com/dbca-wa/wastdr/issues).

```{r export_data}
x %>% wastdr::export_wastd_turtledata(outdir = fn, filename = zipfn)
# readRDS(fs::path(fn, "wastd_data.rds")) to restore x
```

## Data sources

Data for this summary is supplied from WAStD, and is manipulated using
helpers from the R package [wastdr](https://dbca-wa.github.io/wastdr/).
Analysts are welcome to utilise `wastdr` or contribute ideas back as
feature requests.

The raw data is exported into separate CSV files. The content of the
individual CSV files is documented at
[wastdr::export_wastd_turtledata](https://dbca-wa.github.io/wastdr/reference/export_wastd_turtledata.html).

Additionally, an R object of class `wastd_data` containing all the data
objects is saved to `wastd_data.RData` which can be read into R directly.
All code in this workbook can be reproduced by running the code shown
inline with the data from `wastd_data.RData`.

In this report, we filter the data to the location `r params$area_name`.

The data was downloaded from WAStD on `r wastd_data$downloaded_on`.

Training and test data is marked as species "Hatchback turtle (Corolla corolla)",
surveys are "production=FALSE", disturbance causes are "training or test record".
Where summaries hide the columns "species" and "production", respectively, 
training and test data are excluded.

# Turtle nesting

The main caveat here is that some turtle monitoring programs monitor turtle
nesting via the morning track counts, others conduct additional night time 
tagging.

As of now, tagging data is handled in another database, and an export of that
database has to be manually edited into the track count data here.

The report here shows only data from track counts. This includes the missed
tracks during turtle tagging.

This means:

* Total turtle numbers exclude tagged turtles (if there was tagging).
* Nesting success, the fraction of successful (nesting) emergences vs all 
  (including non-nesting emergences), is skewed by the exclusion of tagged 
  turtles (which could have nested or not before being tagged).
  
  
This caveat will be resolved once the tagging database is fully synchronised 
with WAStD, and WAStD therefore contains all tagging records, which then are
added to the descriptive statistics shown in this report.

## Turtle Nesting Current Season

```{r tracks_map_2020}
x$tracks %>% 
  wastdr::filter_wastd_season(2020) %>% 
  wastdr::filter_realspecies() %>% 
  wastdr::map_tracks(sites = x$sites)
```

For performance, we include the same map with tracks clustered by species.
On slow machines with large datasets, this map will perform better than the
above map with individual records.

```{r tracks_map_2020_cluster}
x$tracks %>% 
  wastdr::filter_wastd_season(2020) %>% 
  wastdr::filter_realspecies() %>% 
  wastdr::map_tracks(sites = x$sites, cluster = TRUE)
```


## QA: tracks outside of known sites
```{r tracks_map_nosite}
tracks_nosite <- x$tracks %>% 
  wastdr::filter_realspecies() %>% 
  dplyr::filter(is.na(site_name)) %>% 
  dplyr::mutate(wastd=glue::glue("https://wastd.dbca.wa.gov.au{absolute_admin_url}"))

tracks_nosite %>% wastdr::map_tracks(sites = x$sites)
tracks_nosite %>% rt()
```

## Missed tracks

This section is relevant for monitoring programs which tag turtles.

Missed tracks come from turtles which got away just before or during
tagging. These records can be used to inform tagging start and end
times:

Missed tracks recorded at the beginning of a tagging night means that
tagging needs to start earlier in the day. Missed tracks recorded during
tagging indicates that tagging beaches are understaffed.

Missed tracks are contained in the exported `tracks.csv` with `nest_age` "missed".

```{r tracks_missed}
mt <- x$tracks %>% 
  wastdr::filter_wastd_season(2020) %>% 
  wastdr::filter_realspecies() %>% 
  dplyr::filter(nest_age == "missed")

# mt %>% rt
mt %>% wastdr::map_tracks(sites = x$sites)
```

## Disturbance, predation: nests and general
General disturbance and signs of predator presence is captured through the 
ODK form "Predator or Disturbance". Disturbance or predation of nests are 
captured through the ODK form "Turtle Track or Nest".
As the data specific to the disturbance or predation is identical, both are 
presented here as one table.

In the exported `nest_dist.csv`, the `encounter_type` "nest" refers to disturbances
or predation of turtle nests, whereas "other" refers to sightings of general
disturbance or presence of predators.

```{r dist_map_2020}
x$nest_dist %>% 
  wastdr::filter_wastd_season(2020) %>%
  wastdr::map_dist()
```


### Disturbance causes
The following interactive table tallies both nest and general ("other") 
disturbance by season. Filter this table to inspect disturbance causes.

You can also select any combination of disturbance causes in the map above to 
inspect their spatial distribution across the studied area.

The numbers shown are absolute numbers and not normalised for survey effort or
extent of surveyed area.

```{r dist_causes}
x$nest_dist %>% 
  wastdr::disturbance_by_season() %>% 
  dplyr::arrange(-season, -n) %>% 
  # dplyr::filter(n>1) %>% 
  rt()
```

## Turtle tracks and nests season totals

Tracks and nests are tallied by type and grouped by

-   season: numeric year of season start, i.e. 2019 = season 2019-20

-   nest age:

    -   fresh - made last night,
    -   old - made before last night,
    -   missed - at night during tagging
    
-   species: "cheloniidae-fam" = species unknown

To see only missed tracks, type "mis" into the filter box of column
"(details\_)nest_age".

This summary is exported as `nesting_type_by_season_age_species.csv`.

```{r tracks_season, fig.width=10}
wastd_nesting_type_by_season_age_species <-
  x$tracks %>% 
  wastdr::filter_realspecies() %>% 
  wastdr::nesting_type_by_area_season_age_species() %>% 
  dplyr::arrange(-season)

wastd_nesting_type_by_season_age_species %>%
  readr::write_csv(here::here(fn, "nesting_type_by_season_age_species.csv"))
  
wastd_nesting_type_by_season_age_species %>% 
  # janitor::clean_names(case="sentence") 
  dplyr::select(
    area_name,
    season,
    species,
    nest_age,
    dplyr::contains("-"),
    dplyr::contains("nest")
  ) %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Nesting season totals"
  ) %>% 
  gt::tab_header(
    title="Seasonal nesting totals",
    subtitle="Nesting by rookery, species, nest age tallied by nesting success"
  ) %>%
  gt::cols_label(
    area_name = "Rookery",
    season = "Season",
    species = "Species",
    nest_age = "Nest age"
  )

```

<!--
# `successful-crawl` = "Track with nest",
# `false-crawl` = "Track without nest",
# `track-not-assessed` = "Track (not assessed)",
# `track-unsure` = "Track (unsure)",
# `nest` = "Nest (incubating)",
# `hatched-nest` = "Nest (hatched)",
# `body-pit` = "Body pit") %>% {
#     
#     if (
#       "successful-crawl" %in% names(wastd_nesting_type_by_season_age_species)
#     ){
#       gt::tab_spanner(
#         label = "Successful nesting",
#         columns = c(`successful-crawl`)
#       )} else .
#   } 
# %>% 
#   gt::tab_spanner(
#     label = "Unsuccessful or unconfirmed nesting",
#     columns = c(
#       `false-crawl`, 
#       `track-not-assessed`,
#       `track-unsure`,
#       `nest`,
#       `hatched-nest`,
#       `body-pit`)
#   )
-->

## Turtle tracks and nests daily totals

The same data, but grouped by day instead of season.
This summary is exported as `nesting_type_by_day_age_species.csv`.

The data are shown as an interactive table with search, filtering, and sorting
enabled. The table can be used to inspect nesting activity on individual days.

```{r tracks_day}
wastd_nesting_type_by_day_age_species <-
  x$tracks %>% 
  wastdr::filter_realspecies() %>% 
  wastdr::nesting_type_by_season_calendarday_age_species() %>% 
  dplyr::arrange(-season)

wastd_nesting_type_by_day_age_species %>%
  readr::write_csv(here::here(fn, "nesting_type_by_day_age_species.csv"))
  
wastd_nesting_type_by_day_age_species %>% rt
```

## Night recaptures by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual new turtles by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual recaptures by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual total turtles (individuals) by species year and total

This section will be completed once the tagging database has been merged into WAStD.

## Annual total turtle emergences
Fresh tracks and nests are combined and tallied, grouped by season and species
to form the total number of emergences in one season. 

Older signs of nesting, such as incubating and hatched nests, as well as 
body pits are omitted from the tally.

Caveats: 

* This data needs to be added to the number of tagged turtles from the 
  tagging database.
* The data shown here are not normalised by survey effort nor length of 
  surveyed coastline. 
* Comparison across seasons will be influenced by different survey effort 
  between seasons.
* Comparison to other localities will be influenced by different survey effort 
  between seasons and different length of surveyed coastline between localities.

```{r total_emerg}
x$tracks %>% 
  wastdr::exclude_training_species() %>% 
  dplyr::filter(
    nest_type %in% c(
      "successful-crawl",
      "false-crawl",
      "track-not-assessed",
      "track-unsure"
    )
  ) %>% 
  dplyr::group_by(season, species) %>%
  dplyr::tally() %>%
  dplyr::ungroup() %>%
  dplyr::arrange(-season, -n) %>%
  janitor::clean_names(case="sentence") %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Total turtle emergences"
  ) %>% 
  gt::tab_header(
    title="Seasonal emergence totals",
    subtitle="Emergences tallied by rookery, season, and species"
  ) %>%
  gt::cols_label(
    N = "Emergences"
  ) 
```

## Number of nests tagged

Tagged nests are exported as individual records of encounters with
tagged nests (including initial tagging and re-sighting). 

We show the season tallies, and export the individual records as 
`nest_tags.csv`.

```{r nests_tagged}
wastd_tagged_nests <- x$nest_tags %>% 
  wastdr::exclude_training_species()
```

```{r nests_tagged_season}
wastd_tagged_nests %>% 
  dplyr::group_by(season, species) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season) %>%
  janitor::clean_names(case="sentence") %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Total nest tags"
  ) %>% 
  gt::tab_header(
    title="Total nest tags",
    subtitle="Tally includes any encounter of a nest tag: deployment, re-sighting, and recovery"
  ) %>%
  gt::cols_label(
    N = "Tag encounters"
  ) 
```

## Number of nests found

The data here is the same as above, nest type "successful-crawl"
(confirmed fresh nest) and "hatched nest".

This summary is exported as `fresh_and_hatched_nests.csv`.

```{r nests_total}
wastd_nests <- x$tracks %>% 
  wastdr::exclude_training_species() %>% 
  dplyr::filter(nest_type %in% c("successful-crawl", "hatched-nest")) %>% 
  dplyr::group_by(season, species, nest_type) %>% 
  dplyr::tally() %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season)

wastd_nests %>%
  readr::write_csv(here::here(fn, "fresh_and_hatched_nests.csv"))

wastd_nests %>% 
  dplyr::arrange(-season, -n) %>% 
  # dplyr::mutate(species = wastdr::humanize(species)) %>%
  janitor::clean_names(case="sentence") %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Total nests found"
  ) %>% 
  gt::tab_header(
    title="Total nests found",
    subtitle="Nest encounters grouped by season, species, and nest type"
  ) %>%
  gt::cols_label(
    N = "Nest encounters"
  ) 
```

## Nest loggers

Nest loggers include both initial deployment and any recorded re-sighting.
Multiple loggers can be associated with one nest.

For want of a better overview, we show an interactive table of all loggers here.
The table supports search, filtering, and ordering.

The nest logger records are also exported as `nest_loggers.csv`.

```{r nest_log}
x$nest_loggers %>% rt
```

## Hatching and emergence success
Nest excavations are shown here as statistics grouped by season and species and 
exported as individual nest excavation records. 

This summary is exported as `nest_excavations_hatching_emergence_success.csv`.

```{r nest_excav}
nest_excav <- x$nest_excavations %>%
  wastdr::hatching_emergence_success() %>% 
  dplyr::arrange(-season)

nest_excav %>% 
  readr::write_csv(
    here::here(fn, "nest_excavations_hatching_emergence_success.csv"))

nest_excav %>% 
  dplyr::arrange(-season, species) %>% 
  dplyr::mutate(species = janitor::make_clean_names(species, case="sentence")) %>% 
  janitor::clean_names(case="sentence") %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Hatching and emergence success"
  ) %>% 
  gt::tab_header(
    title="Hatching and emergence success",
    subtitle="Descriptive statistics from all excavated nests"
  )
```

<!-- TODO Plot HS/ES time series to show different season length / survey effort. -->

## Fan angle data

Fan angle data is exported as `nest_fans.csv` and previewed in the following table.
More meaningful visualisations are in preparation.

```{r nest_fans}
x$nest_fans %>% 
  janitor::clean_names(case="sentence") %>% 
  rt
```

## Tag return crossovers from or to by year

This section will be completed once the tagging database has been merged into WAStD.

# Marine Wildlife Incidents

MWI include strandings, rescues, and live sightings of turtles and other 
marine megafauna.

## Live sightings: rescues, in water, tagging
The ODK form "Turtle Sighting" captures turtles encountered during boat based 
surveys.
The ODK form "Marine Wildlife Incident" also captures encounters with injured,
entangled, rescued, but overall alive animals (including turtles).
The ODK form "Turtle Tagging", as well as data import from the legacy turtle
tagging database WAMTRAM, will show up as a live sighting with a nesting
activity.

All of the above forms feed into WAStD AnimalEncounters, which are shown in the map below.

```{r mwi_map_resc}
x$animals %>% 
  wastdr::filter_alive() %>% 
  wastdr::map_mwi(sites = x$sites)
```

## Dead sightings: strandings, mortalities
The ODK form "Marine Wildlife Incident" captures encounters with dead animals 
(including turtles), which then become WAStD AnimalEncounters.

```{r mwi_map_str}
x$animals %>% 
  wastdr::filter_dead() %>% 
  wastdr::map_mwi(sites = x$sites)
```

# Surveys

A summary of past seasons follows. Note the dates are "turtle dates"
(calendar date of nesting night's evening).

This summary is exported as `survey_season_stats.csv`.

```{r svy_stats, fig.width=10}
survey_season_stats <- x$surveys %>% 
  wastdr::filter_realsurveys() %>% 
  wastdr::survey_season_stats() 

survey_season_stats %>% 
  readr::write_csv(here::here(fn, "survey_season_stats.csv"))

survey_season_stats %>% 
  janitor::clean_names(case="sentence") %>% 
  gt::gt(
    groupname_col=NULL,
    caption = "Seasonal overview of survey effort"
  ) %>% 
  # gt::fmt_date(c("First day", "Last day")) %>% # Friday, October 20, 2017
  gt::tab_header(
    title="Seasonal overview of survey effort",
    subtitle="Survey stats grouped by rookery and season"
  )
```

The following summary splits up the above table by sites 
and is exported as `survey_season_site_stats.csv`.

The table is interactive and supports search, filtering and ordering.

```{r svy_stats_site, fig.width=10}
survey_season_site_stats <- x$surveys %>% 
  wastdr::filter_realsurveys() %>% 
  wastdr::survey_season_site_stats()

survey_season_site_stats %>% 
  readr::write_csv(here::here(fn, "survey_season_site_stats.csv"))

survey_season_site_stats %>% 
  janitor::clean_names(case="sentence") %>% 
  rt
```

### QA: Surveys with start points outside the nominated site

### QA: Surveys with end points outside the nominated site

### QA: Backwards Surveys

Incorrectly reconstructed Surveys can run backwards, i.e. with a start time
past the end time.

Saving a "backwards" Survey in WAStD will automatically adjust the end time to 
6h past the start time, then adopt Encounters within that time span. 
Saving the Survey again will set the end time to 30 min past the last adopted 
Encounter.

After QA, this list should be empty.

```{r}
x$surveys %>%
  dplyr::filter(start_time > end_time) %>%
  dplyr::select(
    site_name,
    reporter,
    start_time,
    end_time,
    is_production,
    id
  ) %>%
  janitor::clean_names(case="sentence") %>% 
  dplyr::mutate(
    Edit = glue::glue(
        "<a href=\"https://wastd.dbca.wa.gov.au/",
        "admin/observations/survey/{Id}/change/\" target=\"_\">Edit</a>"
      ),
    View = glue::glue(
        "<a href=\"https://wastd.dbca.wa.gov.au/",
        "observations/surveys/{Id}\" target=\"_\">View</a>"
      )
  ) %>% 
  rt()
```

### QA: Duplicate morning surveys

The following tables shows days and sites for which more than one 
production (= real, not training or duplicate) survey exists in WAStD.
Typically, only one of them should be a "production" survey, 
and all other surveys could be marked and closed as duplicates.

Morning surveys are expected to start after 05:00 and before 15:00.
More than one morning survey on the same site indicates a possible duplicate.

<!-- https://github.com/dbca-wa/wastdr/issues/54 exclude false positives -->
```{r dup_svy_morning}
morning_survey_start_hour <- 5
night_survey_start_hour <- 15

duplicate_morning_surveys <- x$surveys %>% 
  dplyr::filter(lubridate::hour(start_time) > morning_survey_start_hour,
                lubridate::hour(start_time) < night_survey_start_hour) %>% 
  wastdr::duplicate_surveys()

duplicate_morning_surveys %>% 
  dplyr::arrange(-season, site_name) %>% 
  dplyr::select(-site_id) %>% 
  janitor::clean_names(case="sentence") %>% rt
```

### QA: Duplicate night surveys
Night surveys are expected to start after 15:00 and before 5:00.
More than one night survey on the same site and date indicates a possible duplicate.
```{r dup_svy_night}
duplicate_night_surveys <- x$surveys %>% 
  dplyr::filter(lubridate::hour(start_time) < morning_survey_start_hour,
                lubridate::hour(start_time) > night_survey_start_hour) %>% 
  wastdr::duplicate_surveys()

duplicate_night_surveys %>%
  dplyr::arrange(-season, site_name) %>% 
  dplyr::select(-site_id) %>% 
  janitor::clean_names(case="sentence") %>% rt
```
<!--
  # gt::gt(
  #   groupname_col=NULL,
  #   caption = "Duplicated morning surveys"
  # ) %>% 
  # gt::tab_header("Merge duplicated morning surveys") %>% 
  # gt::cols_label(
  #   season = "Season",
  #   site_name = "Site",
  #   calendar_date_awst = "Date (AWST)",
  #   n = "Duplicates",
  #   wastd = "Review and merge"
  # ) %>% 
  # gt::fmt_markdown(columns = vars(wastd))
-->
## QA: Surveys with reconstructed duration of exactly 6 hours
Manually adjust end times as and if appropriate.

```{r reconstructed_scy}
rec_surveys <- x$surveys %>% 
  dplyr::filter(
    lubridate::interval(
      start = start_time, 
      end = end_time) == lubridate::hours(6))

rec_surveys %>% 
  dplyr::select(
    season,
    site_name,
    reporter,
    calendar_date_awst,
    start_time,
    end_time,
    is_production,
    change_url
  ) %>% 
  janitor::clean_names(case="sentence") %>% rt
```

<!--
 # gt::gt(
 #    groupname_col=NULL,
 #    caption = "Surveys with reconstructed end time"
 #  ) %>% 
 #  gt::tab_header("Adjust reconstructed end time") %>% 
 #  gt::cols_label(
 #    season = "Season",
 #    site_name = "Site",
 #    reporter = "Reporter",
 #    calendar_date_awst = "Date (AWST)",
 #    start_time = "Survey start",
 #    end_time = "Survey end",
 #    is_production = "Production survey",
 #    change_url = "Adjust length"
 #  ) %>% 
 #  gt::fmt_markdown(columns = vars(change_url))
-->

## QA: Suspiciously long surveys
Surveys running for more than 6 hours could have been erroneously merged from 
two separate surveys, or they are truly long surveys.

Fixing these requires individual inspection and full fluency in WAStD's operation.

These can capture encounters from two separate surveys. Survey effort can be
biased by showing only one survey where two have been conducted, and by 
showing the extra hours between actual fieldwork as survey duration.

Protocol:

* Open each "Update survey XXX" link to view the survey in the data curation 
  portal.
* From the data curation portal, "view on site" to inspect duration and the 
  list of encounters.
* Do the encounters come from two different dates? If so, find surveys from the
  other date(s). 
  * Shortcut: Survey detail page > map > click site > surveys, then enter the
    desired date into the very last filter field and filter.
  * Long route: Surveys > list surveys > filter by site and date.
* Having all candidate surveys open in the data curation portal:
  * mark both as "production",
  * adjust start and end times to be one one day only, then 
  * aave and continue editing the "recipient" survey, then refresh the "view" 
    and verify that additional encounters were adopted.
  * Save and continue editing the "donor" survey, refresh the "view" and
    verify that the encounters of the incorrect date are gone from the list of
    encounters.
* Refresh both surveys' "view on site". Are the encounters allocated correctly?
  If not, adjust survey start and end times again to enclose their encounters and
  save the surveys again until the number of linked encounters remains stable.
* If the surveys have been split up to separate (consecutive) dates:
  * they should not appear as each others' duplicates, and 
  * their encounters should also be from the respective date.
  * Each survey can have other duplicate surveys from their respective date.
* Always mark the recipient survey production first, then the donor survey.
  Verify that the number of encounters as well as the start and end dates 
  have not shifted.
* If other duplicate surveys were found, these should also be merged correctly now.



```{r long_scy}
long_surveys <- x$surveys %>% 
  dplyr::filter(
    lubridate::interval(
      start = start_time, 
      end = end_time) > lubridate::hours(6))

long_surveys %>%
  dplyr::select(
    season,
    site_name,
    reporter,
    calendar_date_awst,
    start_time,
    end_time,
    is_production,
    change_url
  ) %>% 
  janitor::clean_names(case="sentence") %>% 
  rt()
```

<!--
 # gt::gt(
 #    groupname_col=NULL,
 #    caption = "Suspiciously long surveys"
 #  ) %>% 
 #  gt::tab_header("Adjust start or end time") %>% 
 #  gt::cols_label(
 #    season = "Season",
 #    site_name = "Site",
 #    reporter = "Reporter",
 #    calendar_date_awst = "Date (AWST)",
 #    start_time = "Survey start",
 #    end_time = "Survey end",
 #    is_production = "Production survey",
 #    change_url = "Adjust times"
 #  ) %>% 
 #  gt::fmt_markdown(columns = c("change_url"))
-->
# Personnel

The number of surveys lead by each volunteer is tallied here for each
season.

If user mapping during data import has matched the wrong WAStD user, this table
offers a link to the "Transfer data" form, where all records against one \
(incorrectly matched) WAStD user can be transferred within a rookery
(pre-filled to `r params$area_name`) to another WAStD user (chosen by the QA 
operator).

```{r volunteers}
x$surveys %>% 
  dplyr::group_by(season, reporter, reporter_id) %>% 
  dplyr::tally(name = "Surveys") %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange(-season) %>%
  dplyr::mutate(
    "Transfer" = glue::glue(
        "<a href=\"https://wastd.dbca.wa.gov.au/users/",
        "{reporter_id}/transfer-at/{area_id}\" target=\"_\">Transfer data</a>"
      ),
  ) %>% 
  dplyr::select(-reporter_id) %>% 
  janitor::clean_names(case="sentence") %>% 
  rt
```

<!-- 
Scott 
## Number of Indigenous employee days

To be inserted.

## List of sat tags applied in last three years

Source: Seaturtle.org or tagging database.

TODO: make photos mandatory during daytime but not at night
TODO: GIS beach 2, Karratha office
TODO: decide on beach 3-5, RI
-->
